Analysis of MSHA Inspections, Violations, and Accident Data
========================================================

_An open and ongoing analysis of the data provided by the [US Mine Safety & Health Administration](http://www.msha.gov/OpenGovernmentData/OGIMSHA.asp). See [github.com/mwfrost/MSHA](https://github.com/mwfrost/MSHA) for details._

```{r}
require(reshape)
require(plyr)
require(ggplot2)
require(lubridate)
require(xtable)
require(scales)

require(rmongodb)
require(RMongo)
require(RJSONIO)

setwd('~/Code/MSHA/')
Sys.setlocale(locale="C")
```
## Load Data

### Load Mine Data
```{r}
mdat <- read.table('./data/msha_source/Mines.TXT', header=T, sep="|", fill=T, as.is=c(1:59),quote="\"")
t(head(mdat,3))

```

This script uses `rmongodb` to write the flat files to mongo collections, and uses `RMongo` to read them.

*rmongodb* 

Write each source table to a mongodb collection

```{r mongo.write.mines}

data.frame.to.mongo <- function(df, db, col, fields=names(df)) {
  mongo <- mongo.create()
  if (mongo.is.connected(mongo)) {
    for(i in 1:nrow(df)) {
        b <- mongo.bson.from.list(df[i,fields])
        #print(b)
        #print('----------------')
        if (!is.null(b)){
          mongo.insert(mongo, paste(db, col, sep='.'), b)
          }
        else {
          print('NULL BSON object')
          print(t(df[i,]))
              }
      }  
  }
}



fields <- setdiff(names(mdat), 'DIRECTIONS_TO_MINE')
# data.frame.to.mongo(mdat, 'msha', 'mines', fields)

```

Now switch to the `RMongo` library and check the data. Some sample Mongo aggregations on the Mines collection, based on these examples: 

http://docs.mongodb.org/manual/reference/sql-aggregation-comparison/

http://docs.mongodb.org/manual/reference/sql-comparison/

Example mongodb search: `db.mines.find({MINE_ID : 1519137})`


```{r check.mines}

mongo <- mongoDbConnect(dbName="msha", host="localhost",port='27017')

output <- dbAggregate(mongo, "mines",
                       c('{ 
                         $match: { NO_EMPLOYEES: {$gt: 0} } 
                         }'
                        ,'{ 
                         $group: { 
                         _id: "$PRIMARY_SIC", 
                         employees : {$sum:"$NO_EMPLOYEES"}
                         } 
                         }'
                      )
                      )

print(output)

output <- lapply(output, fromJSON)

output <- do.call(rbind, output)
output <- data.frame(output)


```

Compare the time it takes to run a simple aggregation in mongo to the same task with ddply()

```{r compare}
mongo <- mongoDbConnect(dbName="msha", host="localhost",port='27017')

system.time(
  dbAggregate(mongo, "mines",
                       c('{ $match: { NO_EMPLOYEES: {$gt: 0} } }'
                        ,'{ $group: { _id: "$PRIMARY_SIC", employees : {$sum:"$NO_EMPLOYEES"}} }'
                      )
                      )
)


 system.time(
  ddply(mdat[mdat$NO_EMPLOYEES > 0,], .(PRIMARY_SIC), summarize, employees=sum(NO_EMPLOYEES))
           )
```

Sample rmongodb queries:

```{r rmongodb.samples}
  mongo1 <- mongo.create()
  mongo.count(mongo1, 'msha.mines', query=mongo.bson.empty())
```





Sample RMongo queries:

```{r rmongo.sample.queries}

dbGetQuery(mongo, 'mines', '{STATE: "WV", "COAL_METAL_IND" :"C"}')

# Robinson Run
dbGetQuery(mongo, 'mines', '{MINE_ID: 4601318}')

dbAggregate(mongo, "mines",
                       c(
                        '{ 
                         $group: { 
                         _id: "$PRIMARY_SIC", 
                         employees : {$sum:"$NO_EMPLOYEES"}
                         } 
                         }'
                      )
                      )

```

Evantually add some utility functions

```{r mongo.functions}


```


### Inspections

Inspections.txt is too long to store in memory in R, so read pieces of it and load those into mongo.

```{r}
skip_count <- 250000
start_row <- skip_count + 1

idat <- read.table('./data/msha_source/Inspections.TXT', nrows=skip_count, header=T, sep="|", fill=T,  quote="\"",comment.char = "")
inames <- names(idat)

while (start_row < 2000000) {
     print(paste("About to scan records ", start_row, " through " , start_row + skip_count))

  idat_temp <- read.table('./data/msha_source/Inspections.TXT', nrows=skip_count, header=F, sep="|", fill=T,  skip = start_row ,quote="\"",comment.char = "")
	names(idat_temp) <- inames
	print(paste("Rows collected: " , nrow(idat)))
     
#  data.frame.to.mongo(idat_temp, 'msha', 'inspections', inames)
   
     
	start_row <- start_row + skip_count
}

  mongo1 <- mongo.create()
  mongo.count(mongo1, 'msha.inspections', query=mongo.bson.empty())
```

Compare the operation time

```{r compare.inspections}

system.time(
  dbAggregate(mongo, "inspections",
                       c('{ $match: { CURRENT_CONTROLLER_NAME: "CONSOL Energy Inc" } }'
                        ,'{ $group: { _id: "$CURRENT_MINE_NAME", inspections : {$sum:1}} }'
                      )
                      )
)

system.time(
  ddply(
    subset(idat, CURRENT_CONTROLLER_NAME == 'CONSOL Energy Inc'), 
    .(CURRENT_MINE_NAME) , nrow
    )
)


```


### Accidents
```{r}
# adat <- read.table('./data/msha_source/Accidents.TXT', header=T, sep="|", fill=T, quote="\"",comment.char = "")

# data.frame.to.mongo(adat, 'msha', 'accidents')

  mongo1 <- mongo.create()
  mongo.count(mongo1, 'msha.accidents', query=mongo.bson.empty())

```

### Violations

```{r}
skip_count <- 250000
start_row <- skip_count + 1

# vdat <- read.table('./data/msha_source/Violations.TXT', nrows=skip_count, header=T, sep="|", fill=T, as.is=c(1:55), quote="\"",comment.char = "")
vnames <- names(vdat)

while (start_row < 2000000) {
     print(paste("About to scan records ", start_row, " through " , start_row + skip_count))

  vdat_temp <- read.table('./data/msha_source/Violations.TXT', nrows=skip_count, header=F, sep="|", fill=T, as.is=c(1:55), skip = start_row ,quote="",comment.char = "")
	names(vdat_temp) <- vnames

#  data.frame.to.mongo(vdat_temp, 'msha', 'violations')
	
  print(paste("Rows collected: " , nrow(vdat)))
#	print(paste("Events ", vdat[last_row,c("EVENT_NO")] , " through " , vdat[nrow(vdat),c("EVENT_NO")]))
	start_row <- start_row + skip_count
#	last_row <- nrow(vdat)
}


```
Check the violations data.

```{r check.violations}
  mongo1 <- mongo.create()
  mongo.count(mongo1, 'msha.violations', query=mongo.bson.empty())
```



Next steps:

- Reduce the data sets to only include mines that aren't abandoned

- For each accident, calculate the days since the previous accident at that mine

```{r since.last.accident}
adat.x <- subset(adat, MINE_ID %in% c(4601318) )
ddply(adat.x, .(MINE_ID), 
```


- Normalize inspections, violations, and accidents into "event" objects.

- Add a time series for events to each mine object

Reference links:

[seanhess]: http://seanhess.github.io/2012/02/01/mongodb_relational.html
[mongodb]: http://docs.mongodb.org/manual/tutorial/query-documents/
[mongodb 2]: http://docs.mongodb.org/manual/reference/sql-comparison/
[mongodb 3]: http://docs.mongodb.org/manual/reference/method/db.collection.find/#db.collection.find
[mongodb 4]: http://docs.mongodb.org/manual/reference/sql-aggregation-comparison/
[mongodb 5]: http://docs.mongodb.org/manual/reference/aggregation/operators/
[mongodb 6]: http://docs.mongodb.org/manual/tutorial/aggregation-examples/
[msha]: http://www.msha.gov/OpenGovernmentData/OGIMSHA.asp
[readthedocs]: https://media.readthedocs.org/pdf/a-little-book-of-r-for-time-series/latest/a-little-book-of-r-for-time-series.pdf

```{r denormalize}

```



