Analysis of MSHA Inspections, Violations, and Accident Data
========================================================

_An open and ongoing analysis of the data provided by the [US Mine Safety & Health Administration](http://www.msha.gov/OpenGovernmentData/OGIMSHA.asp). See [github.com/mwfrost/MSHA](https://github.com/mwfrost/MSHA) for details._

```{r}
require(reshape)
require(plyr)
require(ggplot2)
require(lubridate)
require(xtable)
require(scales)
require(RJSONIO)

setwd('~/Code/MSHA/')
source('./lib/calendarHeat.r')
Sys.setlocale(locale="C")
```

### Utility scripts
(not run in this process)

Download the most recent files
`.\data\msha_source\download_source.sh`

Unzip the files, replacing the old copies
`\data\msha_source\unzip_source.sh\`

Use the python `pandas` library to select and write out files small enough for R to handle gracefully.
`.\extract_wv.py`

```{r read.csv.files}
mdat <- read.csv('data/wv_mines.csv')
vdat <- read.csv('data/wv_violations.csv')
idat <- read.csv('data/wv_inspections.csv')
adat <- read.csv('data/wv_accidents.csv')
```

```{r add.dates}

mdat$current.status.dt <- mdy(as.character(mdat$CURRENT_STATUS_DT))
vdat$violation.occur.dt <- mdy(as.character(vdat$VIOLATION_OCCUR_DT))

idat$inspection.begin.dt <- ymd(gsub('-','/', gsub(' 00:00:00','',idat$INSPECTION_BEGIN_DT)))
adat$accident.dt <- ymd(gsub('-','/', gsub(' 00:00:00','',adat$ACCIDENT_DT)))

```


```{r trailing.violations}

trailing.events <- function(mine, acc.dt=now(),trailing.days=90 ){
      d <- data.frame(mine, acc.dt)
      ddply(d, .(mine, acc.dt), function(x){ 
      c(trailing.accidents=nrow(subset(adat, MINE_ID == x$mine 
                                   & accident.dt < x$acc.dt
                                   & accident.dt > x$acc.dt - ddays(trailing.days)
                                   )),
    trailing.violations=nrow(subset(vdat, MINE_ID == x$mine 
                                   & violation.occur.dt < x$acc.dt
                                   & violation.occur.dt > x$acc.dt - ddays(trailing.days)
                                   )),
    trailing.injuries=sum(subset(vdat, MINE_ID == x$mine 
                                   & violation.occur.dt < x$acc.dt
                                   & violation.occur.dt > x$acc.dt - ddays(trailing.days)
                                   )$NO_INJURIES,na.rm = TRUE),  
    trailing.lost.days=sum(subset(vdat, MINE_ID == x$mine 
                                   & violation.occur.dt < x$acc.dt
                                   & violation.occur.dt > x$acc.dt - ddays(trailing.days)
                                   )$DAYS_LOST,na.rm = TRUE), 
     trailing.inspections=nrow(subset(idat, MINE_ID == x$mine 
                                   & inspection.begin.dt < x$acc.dt
                                   & inspection.begin.dt > x$acc.dt - ddays(trailing.days)
                                   ))       
        )
      }
    )
}

# Tests
# trailing.events(c(4601318, 4601456) , ymd(c('2005-09-08', '2011-08-05')), 30)

# trailing.events(adat[1:10,'MINE_ID'], adat[1:10, 'accident.dt'], 30)

```

Every day of operation for every mine needs a record, so we can try to identify any differences between the days without accidents and those with. We don't have the full span of operating days, though. One way to fake it is to run the span between the first and last recorded inspections for each mine

```{r mine.days}

# x is the data frame of inspections
fill.days <- function(x){
  data.frame(
    mine=x$MINE_ID[1],
    mine.day=seq(min(x$inspection.begin.dt), max(x$inspection.begin.dt), by ="day")
             )
  }
# Tests
# fill.days(idat[idat$MINE_ID==4609060,])

# by(idat[idat$MINE_ID %in% c(4609060,4601318),], idat[idat$MINE_ID %in% c(4609060,4601318),'MINE_ID'], fill.days)

mine.days <- ddply(idat, .(MINE_ID), fill.days) 
mine.days$mine <- NULL


```

Now that there's a unique record for every day of operation (sort of), calculate the trailing accidents, violations, and inspections for each mine/day, then join all the accident records to determine which days involved accidents.

TODO: currenty ignoring the problems caused by multiple accidents per day

TODO: This step could probably be much faster as a pandas aggregation pipeline.

```{r mine.day.stats}
head(trailing.events(mine.days[1:10,'MINE_ID'], mine.days[1:10, 'mine.day'], 30))

mine.day.stats <- trailing.events(mine.days[mine.days$MINE_ID %in% c(4609060,4601318),'MINE_ID'], mine.days[mine.days$MINE_ID %in% c(4609060,4601318), 'mine.day'], 30)
# acc.dt is now op.dt
names(mine.day.stats) <- gsub('acc.dt','op.dt', names(mine.day.stats))

mine.day.stats <- merge(mine.day.stats, adat[,c('MINE_ID','accident.dt','DEGREE_INJURY_CD','CLASSIFICATION')], by.x=c('mine','op.dt'), by.y=c('MINE_ID','accident.dt'), all.x=TRUE)

head(subset(mine.day.stats, !is.na(DEGREE_INJURY_CD)))
```



Join some useful categories to the per mine/day stats and flag accident days

```{r mine.joins}
mine.day.stats <- merge(mine.day.stats, 
                        mdat[
                          ,
                          c('MINE_ID','CURRENT_MINE_NAME','COAL_METAL_IND','CURRENT_OPERATOR_NAME','CURRENT_CONTROLLER_NAME','STATE','FIPS_CNTY_NM','NO_EMPLOYEES')
                             ] ,
                        by.x=c('mine'),
                        by.y=c('MINE_ID')
                        )
mine.day.stats$bin.inj <- ifelse(is.na(mine.day.stats$DEGREE_INJURY_CD), 0, 1)
mine.day.stats$bin.acc <- ifelse(is.na(mine.day.stats$CLASSIFICATION), 0, 1)

head(subset(mine.day.stats, p.acc==1.0))
```

```{r two.mine.plot}
table(mine.day.stats$CLASSIFICATION)

ggplot(mine.day.stats) + 
  geom_segment(data=
                 mine.day.stats[
                   mine.day.stats$bin.acc==1.0 & grepl('FIRE |IGNITION', mine.day.stats$CLASSIFICATION)
                    ,
                    ]
               , aes(x=op.dt,xend=op.dt, y=150, yend=0),color='yellow') +
  geom_line(aes(x=op.dt,y=trailing.violations)) + 
  geom_line(aes(x=op.dt,y=trailing.accidents),color='red') +
  geom_line(aes(x=op.dt,y=trailing.inspections),color='blue') +
  facet_grid(CURRENT_MINE_NAME~.) 

ggplot(mine.day.stats) + 
  geom_point(aes(x=trailing.accidents,y=trailing.violations)) + stat_smooth(aes(x=trailing.accidents,y=trailing.violations))

ggplot(mine.day.stats[mine.day.stats$mine==4601318, ]) + geom_line(aes(x=op.dt, y=trailing.violations), color='red') 

```

```{r logit}
logit.acc <- glm(bin.acc ~ trailing.accidents + trailing.violations + trailing.inspections, data = mine.day.stats, family = "binomial")
summary(logit.acc)
confint(logit.acc)

```
