Analysis of panda-provided Modeled Data
========================================================


```{r}
require(reshape)
require(plyr)
require(ggplot2)
require(lubridate)
require(xtable)
require(scales)
require(glmnet)
require(survival)
require(TestSurvRec)

setwd('~/Code/MSHA/')
source('./lib/calendarHeat.r')
Sys.setlocale(locale="C")
```

```{r read.csv.files}

mines <-  read.csv('data/wv_mines.csv', stringsAsFactors=FALSE)
op_days <- read.csv('data/iav_days.csv', stringsAsFactors=FALSE)

op_days$op_date <- gsub('-','/', gsub(' 00:00:00','',op_days$OP_DT))
op_days$op_date <- ymd(op_days$op_date)

op_days$op_year <- year(op_days$op_date)
op_days$op_month <- month(op_days$op_date)
op_days$op_day <- day(op_days$op_date)

op_days$fakeday <- ymd(paste('2012',op_days$op_month,op_days$op_day,sep='/'))

```

Pick a few mines and plot their profiles

```{r sample.plots}
# Use mines with abundant data
sample.mines <- subset(ddply(op_days, .(MINE_ID), summarize, vsum=sum(vcount)), vsum > 1000)$MINE_ID

sample.mines <- sample(sample.mines, 3)
sample.mines

sample.dat <- op_days[op_days$MINE_ID %in% sample.mines & op_days$op_year>2009,]
ggplot(sample.dat, aes(x=fakeday)) + 
    geom_line(aes( y=a30), color='red') + 
    geom_line(aes(y=a90), color='red', linetype='dashed') + 
    geom_line(aes(y=i30), color='green')+
    geom_line(aes(y=v30), color='orange')+
  facet_grid(op_year~MINE_ID)

```




Trying the `glm` library's logit model

```{r logit}
logit.acc <- glm(has_acc ~ a30 + v30 + i30, data = op_days, family = "binomial")
summary(logit.acc)
confint(logit.acc)

mine.day.stats$pred.acc <- predict(logit.acc, op_days[,c('v30','a30')])

ggplot(op_days) + geom_point(aes(x=v30, y=pred.acc, color=a30))

```





Join some useful categories to the per mine/day stats and flag accident days

```{r mine.joins}
mine.day.stats <- merge(mine.day.stats, 
                        mdat[
                          ,
                          c('MINE_ID','CURRENT_MINE_NAME','COAL_METAL_IND','CURRENT_OPERATOR_NAME','CURRENT_CONTROLLER_NAME','STATE','FIPS_CNTY_NM','NO_EMPLOYEES')
                             ] ,
                        by.x=c('mine'),
                        by.y=c('MINE_ID')
                        )
mine.day.stats$bin.inj <- ifelse(is.na(mine.day.stats$DEGREE_INJURY_CD), 0, 1)
mine.day.stats$bin.acc <- ifelse(is.na(mine.day.stats$CLASSIFICATION), 0, 1)

head(subset(mine.day.stats, p.acc==1.0))
```

```{r two.mine.plot}
table(mine.day.stats$CLASSIFICATION)

ggplot(mine.day.stats) + 
  geom_segment(data=
                 mine.day.stats[
                   mine.day.stats$bin.acc==1.0 & grepl('FIRE |IGNITION', mine.day.stats$CLASSIFICATION)
                    ,
                    ]
               , aes(x=op.dt,xend=op.dt, y=150, yend=0),color='yellow') +
  geom_line(aes(x=op.dt,y=trailing.violations)) + 
  geom_line(aes(x=op.dt,y=trailing.accidents),color='red') +
  geom_line(aes(x=op.dt,y=trailing.inspections),color='blue') +
  facet_grid(CURRENT_MINE_NAME~.) 

ggplot(mine.day.stats) + 
  geom_point(aes(x=trailing.accidents,y=trailing.violations)) + stat_smooth(aes(x=trailing.accidents,y=trailing.violations))

ggplot(mine.day.stats[mine.day.stats$mine==4601318, ]) + geom_line(aes(x=op.dt, y=trailing.violations), color='red') 

```


Survival analysis approach

Data format explained [here](http://stats.stackexchange.com/questions/59636/how-do-you-prepare-longitudinal-data-for-survival-analysis)
Instead of day-by-day records, calculate accident-free periods with start and end dates, and assign a status of 1 for each accident

```{r survival}
# transform accident records into periods
# each accident is the end date of a period, and the preceding accident is the maximum date less than it

periods <-data.frame(
      MINE_ID=adat$MINE_ID, 
      tstop =adat$accident.dt)

periods <- ddply(periods, .(MINE_ID, tstop), function(x) c(status=1, tstart=max(periods[periods$MINE_ID==x$MINE_ID & periods$tstop < x$tstop,'tstop'])))

periods$time <- as.numeric(periods$tstop - periods$tstart)/(3600*24)
periods$time <- ifelse(IsInf(periods$time), NULL, periods$time)
periods$time <- ifelse(is.infinite(periods$time), NA, periods$time)
periods$id <- periods$MINE_ID

plot(survfit(Surv(time, status) ~ 1, data=periods[periods$id %in% periods[1:20,'id'],] ), 
         conf.int=FALSE, mark.time=FALSE)

```

